Measurement of deeply virtual Compton scattering off Helium-4 with CLAS at Jefferson Lab

### 2020-04-22 eg6-letters/long-paper/prc-dvcshelium.tex
Add Stepan's comment and some fig rework

In nuclei, the GPDs can be probed conveniently through the measurement of the
spin asymmetries generated by the deeply virtual Compton scattering (DVCS) 
process~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya}. The measurement of the exclusive
production of a photon limits the possibilities of final state interactions in the nuclear 
medium and offers a unique opportunity to make a measurement free of final state interactions 
(FSI). Moreover, with a nuclear targets of spin-0 the extraction of the GPD from the DVCS data 
is significantly simplified with the reduction to a single GPD intervening in the process at leading order. The 
measurement of the nuclear DVCS is however challenging experimentally, first attempts from 
the HERMES collaboration \cite{Airapetian:2009cga} lead to controversial conclusions. We present here
in detail the more recent measurements from the CLAS collaboration, which has been already partially presented
in two short letters \cite{Hattawy:2017woc,Hattawy:2018liu}. We extend in this article 
the description of the nuclear DVCS CLAS experiment, detail the methods used for the
data analysis and produce the complete experimental results for each channel measured.

@@

The theory of GPDs has been already reviewed in details in various 
publications~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya},
we summarized here only necessary elements to discuss our experimental results.
The GPDs are real structure functions $F^{q}(x,\xi,t)$, where $x+\xi$ and $x-\xi$ are incoming and outgoing quark momenta 
respectively and $t=\Delta^2$ the squared transferred momentum to the target in this exchange of quarks as illustrated in 
Fig. \ref{fig:GPD}.

@@

DVCS is the main experimental access to GPDs. However, this process does not offer
a full access to the GPDs and we only access them integrated over $x$, in the following form:
\begin{equation}
	\int^1 _{-1} dx F^{q}(x,\xi,t) \left [ {1 \over x - \xi - i \epsilon} 
	                                   \mp {1 \over x + \xi - i \epsilon} \right]. 
\end{equation}
To account for this and simplify equations, we define the complex Compton form factors (CFF,
noted with curved $\mathcal{F}$ for a given GPD $F$) for each GPD as follows:
\begin{equation}
	\Re e (\mathcal{F}(\xi,t)) = \sum _q e^2_q \mathcal{P} \int^1_{-1} dx F^{q}(x,\xi,t)
	      \left [ {1 \over x - \xi} \mp {1 \over x + \xi} \right], 
\end{equation}
\begin{equation}
	\Im m (\mathcal{F}(\xi,t)) = - \pi \sum _q e^2_q \left [ F^{q}(\xi,\xi,t) \mp F^{q}(-\xi,\xi,t) \right]. 
\end{equation} 
These are the quantities directly present in the DVCS cross sections. We note that they are summed over
the different quark flavors present in the hadron, as the electromagnetic probe do not differentiate
quark flavors.

Experimentally, another process is indistinguishable from the DVCS, the Bethe-Heitler (BH) 
process in which 
the final state photon is emitted by the scattering lepton rather than the hadron. In this case, the 
photon-hadron interaction is the same as in elastic scattering and depends on the
target FFs rather than its GPDs. The DVCS and BH processes are experimentally indistinguishable as they
have identical final states, such that they interfere in the squared amplitude of the
exclusive photo-production process:
\begin{equation}
	|T|^2 = |T_{DVCS}|^2 + |T_{BH}|^2 + T_{DVCS}^* T_{BH} + T_{DVCS}T_{BH}^*. 
\end{equation}
The interference terms increase significantly the cross section in specific parts of the phase space
and lead to significant beam spin asymmetries (BSA), which are the focus of the measurements presented 
here.

Finally, we need to define our kinematics properly. We use the 
conventions from Fig. \ref{fig:PhiAngle} for angles and the experimental kinematic variables
used here are defined as: 
$-t = -(p_p-p_p^\prime)^2 = \Delta^2$ and 
$x_B = {Q^2 \over 2 M_N \nu}\sim {2\xi \over 1+\xi}$, with $M_N$ the nucleon mass and $\nu$ the
energy transfer to the target $\nu = {E -E^\prime} $.

@@

The first reaction measured in the experiment is the coherent electro-production of a photon on helium
$e+^4\!\!He \rightarrow e+\gamma+^4\!\!He$ at high 4-momentum transfer squared ($Q^2$). The 
leading order diagram of the nuclear coherent DVCS is represented in Fig. \ref{fig:CohDiag}.
In our experiment, we focused on the measurement of the beam-spin asymmetry (BSA) noted $A_{LU}$ with 
$L$ for the longitudinally-polarized electron beam and $U$ the unpolarized target and defined as:  
\begin{equation}
A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
              {d^{5}\sigma^{+} + d^{5}\sigma^{-}},
  \label{eq:BSA}
\end{equation}
where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the differential cross section for a positive 
(negative) beam helicity. At leading order and leading twist, the BSA can be expressed as \cite{Kirchner:2003wt}:        
\begin{eqnarray}
\label{eq:coh_BSA}
A_{LU}& =& \frac{x_A(1+\epsilon^2)^2}{y} \, s_1^{INT} \sin(\phi) \, 
\bigg/ \, \bigg[ \, \sum_{n=0}^{n=2}c_n^{BH}\cos{(n\phi)} +  \\
& & \frac{x_A^2 t {(1+\epsilon^2)}^2}{Q^2} P_1(\phi) P_2(\phi) \, c_0^{DVCS} + 
\frac{x_A (1+\epsilon^2)^2}{y} \sum_{n=0}^{n=1} c_n^{INT} \cos{(n\phi)} \bigg].  \nonumber 
\end{eqnarray}
where $\mathcal{P}_1(\phi)$ and $\mathcal {P}_2(\phi)$ are the BH 
propagators. The factors: $c_{0,1,2}^{BH}$, $c_0^{DVCS}$, $c_{0,1}^{INT}$ and 
$s_1^{INT}$ are the Fourier coefficients of the BH, the DVCS and the 
interference amplitudes for a spin-zero target. The explicit 
expressions of these coefficients, which have been derived based on the work of 
Kirchner and Müller \cite{Kirchner:2003wt}, can be found in Appendix \ref{sec:eq}.

@@

The incoherent nuclear DVCS process, is the DVCS off a bound nucleon in a nucleus
as represented in Fig. \ref{fig:InCohDiag} for an helium-4 target. The remnants of 
the nucleus ($X$) contain only the missing three nucleons. 
The theory for incoherent DVCS on the nucleon is largely based on the free proton theory
already reviewed many times in the literature~\cite{Diehl:2003ny,Belitsky:2005qn,Guidal:2013rya}. 
Two important differences need to be accounted for however, the different initial state and the addition of 
final state effects. In the initial state, the intrinsic Fermi motion of the nucleons in the nucleus 
lead to an uncertainty on the exact kinematic of the reaction. Moreover, in general the nucleon 
is in an off-shell state that is not exactly identical to its final state. In the final state, 
interactions between the outgoing nucleon from the DVCS reaction and remnants of the nuclear 
target are possible. The latter leading to contamination from other channels, in particular 
charge exchange processes can lead to a large contribution from such background reaction.

@@

Since, DVCS is a 
process selected using tight exclusivity constrains, some of the initial and final-state effects are
automatically mitigated. Selection criterion on missing energy and momentum are performed,
constraining the range of initial Fermi motion and final interactions possible. However, no theoretical
calculation is available to correct for the reminder of these effects yet. Since, modern calculations exist 
for such effects in DIS \cite{Cosyn:2017ekf}, we can expect them to be extended to the DVCS 
process as more data becomes available. Another avenue of progress
on this topic will be the use of experimental techniques like tagging. This process
can help to control both initial and final state effects by detecting the nuclear remnant.
In the tagged process the target breaks in two providing directly the initial state
of the struck nucleon, while a measured backward nuclear fragment limits significantly the
impact of final state effects.

@@

An important issue with the HERMES measurement and how it is obtained from data has been raised in
\cite{Guzey:2003jh} to explain the discrepancy with theoretical expectations.
The main concern is that the DVCS process is not fully detected and the scattered target
is instead reconstructed through a missing mass measurement of the other reaction products. The 
issue with this method is that the detector resolution is not good enough to separate the 
coherent and incoherent channels properly. 
Instead, the results are labeled "coherent enriched" and "incoherent enriched" at low and high 
$-t$, respectively. This label is based on the assumption that the very different behavior of the
cross sections of the two channels in $t$ will lead to a clear differentiation. However, the
results in Fig. \ref{fig:HERMES2} show similar behaviors in both sectors of $t$, which is 
challenging this assumption and could explain the tension between theory and experiment.

Altogether, the large error bars and the impossibility to properly separate the coherent and 
incoherent channels have strongly impaired the measurement and any conclusions that could be
obtained from it. The CLAS experiment presented here has profited largely from this
result and was designed specifically to solve these two issues of low statistics and exclusivity.

@@

The CLAS experiment had for main objective to explore the coherent DVCS on helium-4 and assess if
the predicted BSA increase could be observed and if GPDs could be extracted from it. In order to
perform this measurement however, several instrumentation challenges needed to be resolved. First, to
measure the scattered electron and the low angle photon of the DVCS, we used CLAS in its 
DVCS setup, $i.e.$ with the addition of a low angle calorimeter and a 5-T solenoid magnet. Second, a
radial time projection chamber (RTPC) was installed to measure the
helium recoils and thus ensure the exclusivity of the process in the coherent channel. In this section, we 
will review the important elements of this detection setup.

@@

The CLAS \cite{Mecking:2003zu} is installed in the hall-B of the JLab continuous electron beam 
accelerator facility (CEBAF). CLAS is specifically designed to study the multi-particles final states
that cannot be observed conveniently with arm spectrometers. Its historic use to measure DVCS in many different 
configurations made it an ideal place for this new DVCS measurement. CLAS is composed of six identical radial 
sectors separated by toroidal magnets, with each sector made of four detectors 
as shown in Fig. \ref{fig:CLAS}.  Three regions of drift chambers \cite{Mestayer:2000we} 
are placed in the solenoid area to reconstruct the charge particle's tracks and 
calculate their momentum. Then, an array of scintillators is placed to measure 
a precise time for each track and perform time-of-flight measurement 
\cite{Smith:1999ii}. These detectors cover the polar angle from 8 to 142 degrees. 
They are completed by two important forward detectors, covering the polar angle from 8 
to 45 degrees, a Cerenkov counter 
\cite{Adams:2001kk} and an electromagnetic calorimeter \cite{Amarian:2001zs}.

@@

Altogether, CLAS provides a large acceptance of about $\pi$ steradians for momentums 
starting at 200 MeV/c. During the nuclear DVCS experiment,
the beam energy was maximal for CLAS with 6.064 GeV, while the beam intensity varied
between 120 and 150 nA. Such beam on our helium-4 target pressurized between 5 and 6 atm,
corresponds to luminosities in the range of $1$ to $1.2 \times 10^{34}$ cm$^{-2}$.s$^{-1}$.
During the experiment, the data acquisition rate ran around 3 kHz with about 70\% live-time
using an inclusive electron trigger.

@@

The CLAS collaboration has established a specific setup to measure the typically 
low angle photons of the DVCS process. This setup is composed of an inner 
calorimeter and a solenoid and has lead to numerous DVCS measurements on proton 
targets \cite{Seder:2014cdc,Jo:2015ema,HirlingerSaylor:2018bnu}.

The inner calorimeter, illustrated in Fig. \ref{fig:IC} is a homogeneous 
calorimeter composed 424 lead tungstate 
(PbWO) crystals read out by $5 \times 5$ mm$^2$ avalanche photo-diodes (APDs). 
It covers angles from 4 to 15 degrees. However, placing a detector at such low angles makes it 
particularly sensitive to the low energy Moller electrons scattered from the target.
To protect the calorimeter from this background, a 5~T solenoid is 
placed around the target to form
a magnetic shield. Thanks to this field, low energy charged particles (particularly 
electrons) curl around the beam line 
and never make it to the calorimeter or other CLAS detectors as illustrated 
by simulation results presented 
in Fig. \ref{fig:Solenoid}. This allows to run much higher luminosity experiments,
a necessity for low rates processes like DVCS.

@@

The recoil helium nuclei from coherent DVCS are mostly emitted between 150 
and 200 MeV/c at our beam energy. Therefore, a specific detector was 
needed to detect them. To design the present setup, large inspiration was drawn from the 
Bonus setup that also used a GEM based RTPC~\cite{Fenker:2008zz} in CLAS to 
detect slow protons out of a deuterium target~\cite{Baillie:2011za}. In such an RTPC the 
charges are projected toward large radii 
rather than towards the endcaps, as is more traditional in time 
projection chambers. This design allows to reduce significantly the drift time and reduce the 
amount of pile-up from accidental events. The RTPC design, its operation, calibration and 
the track reconstruction have been described in more 
details elsewhere~\cite{Dupre:2017upj}, here is a summary of key elements.

@@

To select exclusive events and suppress the backgrounds, we define several 
exclusivity cuts based on energy and momentum conservation. In principle, 
a selection based on two or three variables can be used to guaranty the 
exclusivity of the process. However, in such experiment, where particles 
are detected at very different energies and with very different detector 
resolutions, we chose to over constrain the selection by using seven 
variables. The seven variables are defined as follow for the coherent 
DVCS (replace helium by proton for the incoherent case):
\begin{itemize}
	\item Co-planarity ($\Delta \phi$) of the virtual photon, the real photon and
		the recoil helium, defined as the angle between the helium, virtual photon
		plane and the virtual photon, real photon plane;
	\item Missing energy of the full system;
	\item Missing mass of the full system;
	\item Missing transverse momentum of the full system;
	\item Missing mass of the system electron-helium, ignoring the photon;
	\item Missing mass of the system electron-photon, ignoring the recoil helium;
	\item Co-linearity ($\theta$) of the measured photon with the missing momentum of the 
		electron-helium system.
\end{itemize}

In the analysis, we apply selection cuts based on a fit of the exclusive peak at 3$\sigma$ around 
the mean value for each variable. This systematic method helps to avoid any bias 
in the selection of the events. The selection of coherent DVCS with these variables is illustrated in 
Fig. \ref{fig:CohExcCuts}. We note on these distributions only few minor anomalies, where the 
distribution have some asymmetries. These are linked with the detector resolutions, that impact
some of the kinematic variables non linearly. The selection of incoherent DVCS is presented in 
Fig. \ref{fig:IncExcCuts}, with two main differences: larger distributions overall and more
offset distributions. The former is largely due to Fermi motion, but simulations have shown
that this effect is not strong enough to reproduce these distribution widths and final 
state interactions must play a role as well. The later is caused by slight detector 
misalignment between CLAS sectors and are within the levels obtained with free proton 
targets~\cite{HirlingerSaylor:2018bnu} to which they can be directly compared.

@@

The main contamination is expected to be 
from the exclusive production of a $\pi^0$, the final state of which is very similar to DVCS with 
only an extra photon. In such an event, if one of the photon is produced at low energy, it is 
easy to confuse this process with single photon production. In order to estimate the contribution from
this channel in our data, we measure the exclusive $\pi^0$ production in the same way as DVCS, with 
a series of exclusivity cuts, completed by a selection cut on the invariant mass of the two photons
to match the $\pi^0$ mass. The events obtained for the coherent and incoherent
channels are shown in Fig. \ref{fig:CohPi0Simul} and \ref{fig:InCohPi0Simul}, respectively. From 
this sample, we elaborated an event generator, which after being processed
in the simulation of our detectors output the red histograms of Fig. \ref{fig:CohPi0Simul} and 
\ref{fig:InCohPi0Simul}. To correct the experimental data, we estimate the number of single photon 
events coming from DVCS with:
\begin{equation}
	N_{1\gamma,\pi^0}^{Exp} = \frac{N_{1\gamma,\pi^0}^{Sim}}{N_{2\gamma,\pi^0}^{Sim}} \times N_{2\gamma,\pi^0}^{Exp},\\
\end{equation}
where $N_{1\gamma,\pi^0}^{Sim}$ is the number of simulated exclusive $\pi^0$ mistaken for DVCS events,
$N_{2\gamma,\pi^0}^{Sim}$ the number of simulated exclusive $\pi^0$ fully reconstructed and $N_{2\gamma,\pi^0}^{Exp}$
the number of experimentally measured exclusive $\pi^0$. This number can then be subtracted to
the experimental measurement of DVCS events ($N_{DVCS}^{Exp}$) to get the corrected result: 
\begin{equation}
	N_{DVCS}^{Corr} = N_{DVCS}^{Exp} - N_{1\gamma,\pi^0}^{Exp}.\\
\end{equation}

@@

The $\pi^0$ contamination was found to be 2 to 4\% in the coherent channel and 3 to 17\% in the incoherent
channel (with variations between bins). To make the correction on the DVCS BSA, we assume that
the exclusive $\pi^0$ production has no such asymmetry. This has been checked with the exclusive 
$\pi^0$ production data, for which no significant level of BSA has been measured.

@@

\subsection{Systematic errors}

To evaluate the systematic error of the measurement, we performed several specialized studies. We 
evaluated the impact of changing the exclusivity selection cuts by varying them from 1 to 5 $\sigma$. 
We evaluated the impact of changing the binning in $\phi$ on the extraction of the beam-spin asymmetry
at 90°. The precision with which we know the beam polarization is estimated based on the spread 
of its measurements during the complete run period. We also checked different methods to make the 
simulation of the exclusive $\pi^0$ production to evaluate the 
possible bias introduced by this correction. As radiative corrections are expected to be small
for our process, we did not apply them, but associated an error of their expected size.
These errors are summarized in Tab. \ref{Table:systematic_uncertainties}, with their respective 
evaluated sizes. They are added quadratically for the total systematic error presented in the results.

@@

>>>>

In nuclei, the GPDs can be probed conveniently through the measurement of the
spin asymmetries generated by the deeply virtual Compton scattering (DVCS) 
process~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya}. The measurement of the exclusive
production of a photon limits the possibilities of final state interactions in the nuclear 
medium and offers a unique opportunity to make a measurement free of final state interactions 
(FSI). Moreover, with a nuclear targets of spin-0 the extraction of the GPD from the DVCS data 
is significantly simplified with the reduction to a single GPD intervening in the process at leading order. The 
measurement of the nuclear DVCS is however challenging experimentally, first attempts from 
the HERMES collaboration \cite{Airapetian:2009cga} lead to controversial conclusions. We present here
in detail the more recent measurements from the CLAS collaboration, which has been already partially presented
in two short letters \cite{Hattawy:2017woc,Hattawy:2018liu}. We extend in this article 
the description of the CLAS nuclear DVCS experiment, detail the methods used for the
data analysis and produce the complete experimental results for each channel measured.

@@

The theory of GPDs has been already reviewed in details in various 
publications~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya},
we summarized here only necessary elements to discuss the present experimental results.
The GPDs are real structure functions $F^{q}(x,\xi,t)$, where $x+\xi$ and $x-\xi$ are incoming and outgoing quark momenta 
respectively and $t=\Delta^2$ the squared transferred momentum to the target in this exchange of quarks as illustrated in 
Fig. \ref{fig:GPD}.

@@

DVCS is the main experimental probe of the GPDs. However, this process does not allow
for an extraction of the GPDs in full phase space parameters. Instead, DVCS gives access
to the GPDs integrated over $x$, in the following form:
\begin{equation}
	\int^1 _{-1} dx F^{q}(x,\xi,t) \left [ {1 \over x - \xi - i \epsilon} 
	                                   \mp {1 \over x + \xi - i \epsilon} \right]. 
\end{equation}
To account for this and simplify equations, we define the complex Compton form factors (CFF,
noted with curved $\mathcal{F}$ for a given GPD $F$) for each GPD as follows:
\begin{equation}
\Re e (\mathcal{F}(\xi,t)) = \sum _q e^2_q \mathcal{P} \int^1_{-1} dx F^{q}(x,\xi,t)
    \left [ {1 \over x - \xi} \mp {1 \over x + \xi} \right], 
\end{equation}
\begin{equation}
\Im m (\mathcal{F}(\xi,t)) = - \pi \sum _q e^2_q \left [ F^{q}(\xi,\xi,t) \mp F^{q}(-\xi,\xi,t) \right]. 
\end{equation} 
These are the quantities directly present in the DVCS cross sections. We note that they are summed over
the different quark flavors present in the hadron, as the electromagnetic probe do not differentiate
quark flavors.

Experimentally, another process is indistinguishable from the DVCS, the Bethe-Heitler (BH) 
process in which 
the final state photon is emitted by the scattering lepton rather than the hadron. In this case, the 
photon-hadron interaction is the same as in elastic scattering and depends on the
target form factors rather than its GPDs. The DVCS and BH processes are experimentally 
indistinguishable as they
have identical final states, such that they interfere in the squared amplitude of the
exclusive photo-production process:
\begin{equation}
|T|^2 = |T_{DVCS}|^2 + |T_{BH}|^2 + T_{DVCS}^* T_{BH} + T_{DVCS}T_{BH}^*. 
\end{equation}
The interference terms increase significantly the cross section in specific parts of the phase space
and lead to significant beam spin asymmetries (BSA), which are the focus of the measurements presented 
here.

Finally, we need to define the kinematics. We use the 
conventions from Fig. \ref{fig:PhiAngle} for angles and the experimental kinematic variables
used here are defined as: 
$-t = -(p_p-p_p^\prime)^2 = \Delta^2$ and 
$x_B = {Q^2 \over 2 M_N \nu}\sim {2\xi \over 1+\xi}$, with $M_N$ the nucleon mass and $\nu$ the
energy transfer to the target $\nu = {E -E^\prime} $.

@@

The first reaction measured in the experiment is the coherent electro-production of a photon on helium
$e+^4\!\!He \rightarrow e+\gamma+^4\!\!He$ at large 4-momentum transfer squared ($Q^2$). The 
leading order diagram of the nuclear coherent DVCS is represented in Fig. \ref{fig:CohDiag}.
In the present experiment, we focused on the measurement of the beam-spin asymmetry (BSA) 
noted $A_{LU}$ with 
$L$ for the longitudinally-polarized electron beam and $U$ the unpolarized target and is defined as:  
\begin{equation}
A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
              {d^{5}\sigma^{+} + d^{5}\sigma^{-}},
  \label{eq:BSA}
\end{equation}
where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the differential cross section for a positive 
(negative) beam helicity. At leading order and leading twist, the BSA can be expressed as \cite{Kirchner:2003wt}:        
\begin{eqnarray}
\label{eq:coh_BSA}
A_{LU}& =& \frac{x_A(1+\epsilon^2)^2}{y} \, s_1^{INT} \sin(\phi) \, 
\bigg/ \, \bigg[ \, \sum_{n=0}^{n=2}c_n^{BH}\cos{(n\phi)} +  \\
& & \frac{x_A^2 t {(1+\epsilon^2)}^2}{Q^2} P_1(\phi) P_2(\phi) \, c_0^{DVCS} + 
\frac{x_A (1+\epsilon^2)^2}{y} \sum_{n=0}^{n=1} c_n^{INT} \cos{(n\phi)} \bigg].  \nonumber 
\end{eqnarray}
where $\mathcal{P}_1(\phi)$ and $\mathcal {P}_2(\phi)$ are the BH 
propagators. The factors: $c_{0,1,2}^{BH}$, $c_0^{DVCS}$, $c_{0,1}^{INT}$ and 
$s_1^{INT}$ are the Fourier coefficients of the BH, the DVCS and the 
interference amplitudes for a spin-zero target. The explicit 
expressions of these coefficients, which have been derived based on the work of 
Kirchner and Müller \cite{Kirchner:2003wt}, can be found in Appendix \ref{sec:eq}.

@@

The incoherent nuclear DVCS process, is the DVCS off a bound nucleon in a nucleus
as represented in Fig. \ref{fig:InCohDiag} for an helium-4 target. The remnants of 
the nucleus ($X$) contain only the missing three nucleons. 
The theory for incoherent DVCS on the nucleon is largely based on the free proton theory
already reviewed widely in the literature~\cite{Diehl:2003ny,Belitsky:2005qn,Guidal:2013rya}. 
Two important differences need to be accounted for however, the different initial state and the addition of 
final state effects. In the initial state, the intrinsic Fermi motion of the nucleons in the nucleus 
lead to an uncertainty on the exact kinematic of the reaction. Moreover, in general the nucleon 
is in an off-shell state that is not exactly identical to its final state. In the final state, 
interactions between the outgoing nucleon from the DVCS reaction and remnants of the nuclear 
target are possible. The latter leading to contamination from other channels, in particular 
charge exchange processes can lead to a large contribution from such background reaction.

@@

Since, DVCS is a 
process selected using tight exclusivity constrains, some of the initial and final-state effects are
automatically mitigated. Selection criterion on missing energy and momentum are performed,
constraining the range of initial Fermi motion and final interactions possible. However, no theoretical
calculation is available to correct for the reminder of these effects yet. Since, modern calculations exist 
for such effects in DIS \cite{Cosyn:2017ekf}, we can expect them to be extended to the DVCS 
process as more data becomes available. Another avenue of progress
on this topic will be the use of experimental techniques like tagging. This process
can help to control both initial and final state effects by detecting the nuclear remnant.
In the tagged process the target breaks in two providing directly the initial state
of the struck nucleon, while measured backward nuclear fragments limit significantly the
impact of final state effects.

@@

An important issue with the HERMES measurement and how it is obtained from data has been raised in
\cite{Guzey:2003jh} to explain the discrepancy with theoretical expectations.
The main concern is that the DVCS process is not fully detected and the scattered target
is instead reconstructed through a missing mass measurement of the other reaction products. The 
issue with this method is that the detector resolution is not good enough to separate the 
coherent and incoherent channels properly. 
Instead, the results are labeled "coherent enriched" and "incoherent enriched" at low and high 
$-t$, respectively. This label is based on the assumption that the very different behavior of the
cross sections of the two channels in $t$ will lead to a clear differentiation. However, the
results in Fig. \ref{fig:HERMES2} show similar behaviors in both sectors of $t$, which  
challenges this assumption and could explain the tension between theory and experiment.

Altogether, large error bars and the failure to properly separate the coherent and 
incoherent channels have strongly impaired the measurement and any conclusions that could be
obtained from it. The CLAS experiment presented here has profited largely from this
result and was designed specifically to solve these two issues of low statistics and exclusivity.

@@

The CLAS experiment had for main objective to explore the coherent DVCS on helium-4 and assess if
the predicted BSA increase could be observed and if GPDs could be extracted from it. In order to
perform this measurement however, several instrumentation challenges needed to be resolved. First, to
measure the scattered electron and the small angle photon of the DVCS, we used CLAS in its 
DVCS setup, $i.e.$ with the addition of a forward angle calorimeter and a 5-T solenoid magnet. Second, a
radial time projection chamber (RTPC) was installed to measure the
helium recoils and thus ensure the exclusivity of the process in the coherent channel. In this section, we 
will review the important elements of this detection setup.

@@

The CLAS \cite{Mecking:2003zu} is installed in the hall-B of the JLab continuous electron beam 
accelerator facility (CEBAF). CLAS is specifically designed to study the multi-particles final states
that cannot be observed conveniently with arm spectrometers. Its historic use to measure DVCS in many different 
configurations made it an ideal place for this new DVCS measurement. CLAS is composed of six identical radial 
sectors separated by coils of toroidal magnets, with each sector made of four detectors 
as shown in Fig. \ref{fig:CLAS}.  Three regions of drift chambers \cite{Mestayer:2000we} 
are placed between the torus magnet to reconstruct the charge particle's tracks and 
calculate their momentum. An array of scintillators is placed behind drift chambers to measure 
a precise time for each track and perform time-of-flight measurement 
\cite{Smith:1999ii}. These detectors cover the polar angle from 8 to 142 degrees. 
In forward region, from 8 to 45 degree, these two detectors are complimented with 
Cerenkov counters \cite{Adams:2001kk} and electromagnetic calorimeters \cite{Amarian:2001zs},
important for electron identification.

@@

Altogether, CLAS provides a large acceptance of about $2\pi$ steradians for momenta 
starting at 200 MeV/c. During the nuclear DVCS experiment,
the beam energy was highest for CLAS with 6.064 GeV, while the beam intensity varied
between 120 and 150 nA. Such beam on the helium-4 target pressurized between 5 and 6 atm,
corresponds to luminosities in the range of $1$ to $1.2 \times 10^{34}$ cm$^{-2}$.s$^{-1}$.
During the experiment, the data acquisition rate ran around 3 kHz with about 70\% live-time
using an inclusive electron trigger.

@@

The CLAS collaboration has established a specific setup to measure the typically 
small angle photons of the DVCS process. This setup is composed of an inner 
calorimeter and a solenoid and has lead to numerous DVCS measurements on proton 
targets \cite{Seder:2014cdc,Jo:2015ema,HirlingerSaylor:2018bnu}.

The inner calorimeter, illustrated in Fig. \ref{fig:IC} is a homogeneous 
calorimeter composed 424 lead tungstate 
(PbWO) crystals read out by $5 \times 5$ mm$^2$ avalanche photo-diodes (APDs). 
It covers angles from 4 to 15 degrees. However, placing a detector at such small angles makes it 
particularly sensitive to the low energy Moller electrons scattered from the target.
To protect the calorimeter from this background, a 5~T solenoid is 
placed around the target to form
a magnetic shield. Thanks to this field, low energy charged particles (particularly 
electrons) curl around the beam line 
and never make it to the calorimeter or other CLAS detectors as illustrated 
by simulation results presented 
in Fig. \ref{fig:Solenoid}. This allows to run much higher luminosity experiments,
a necessity for low rates processes like DVCS.

@@

The recoil helium nuclei from coherent DVCS are mostly emitted between 150 
and 200 MeV/c at the beam energy of 6~GeV. Therefore, a specific detector was 
needed to detect them. To design the present setup, large inspiration was drawn from the 
Bonus setup that also used a GEM based RTPC~\cite{Fenker:2008zz} in CLAS to 
detect slow protons out of a deuterium target~\cite{Baillie:2011za}. In such an RTPC the 
charges are projected toward large radii 
rather than towards the endcaps, as is more traditional in time 
projection chambers. This design allows to reduce significantly the drift time and reduce the 
amount of pile-up from accidental events. The RTPC design, its operation, calibration and 
the track reconstruction have been described in more 
details elsewhere~\cite{Dupre:2017upj}, here is a summary of key elements.

@@

To select exclusive events and suppress the backgrounds, we define several 
exclusivity cuts based on energy and momentum conservation. In principle, 
a selection based on two or three variables can be used to guaranty the 
exclusivity of the process. However, in such experiment, where particles 
are detected at very different energies and with very different detector 
resolutions, we chose to over constrain the selection by using seven 
variables. The seven variables are defined as follow for the coherent 
DVCS (replace helium by proton for the incoherent case):
\begin{itemize}
	\item Co-planarity ($\Delta \phi$) of the virtual photon, the real photon and
		the recoil helium, defined as the angle between the helium, virtual photon
		plane and the virtual photon, real photon plane;
	\item Missing energy of the full system;
	\item Missing mass of the full system;
	\item Missing transverse momentum of the full system;
	\item Missing mass of the system electron-helium;
	\item Missing mass of the system electron-photon;
	\item Co-linearity ($\theta$) of the measured photon with the missing momentum of the 
		electron-helium system.
\end{itemize}

In the analysis, we apply selection cuts based on a fit of the exclusive peak at 3$\sigma$ around 
the mean value for each variable. This systematic method helps to avoid any bias 
in the selection of the events. The selection of coherent DVCS with these variables is illustrated in 
Fig. \ref{fig:CohExcCuts}. We note on these distributions only few minor anomalies, where the 
distribution have some asymmetries. These are linked with the detector resolutions, that impact
some of the kinematic variables non linearly. The selection of incoherent DVCS is presented in 
Fig. \ref{fig:IncExcCuts}, with two main differences: larger distributions overall and more
offset distributions. The larger distributions are largely attributed to the effect of 
Fermi motion, but simulations have shown that this effect is not strong enough to fully 
reproduce the distribution widths and final state interactions must play a role as well. 
The offsets of some distributions is caused by slight detector 
misalignment between CLAS sectors and are within the levels obtained with free proton 
targets~\cite{HirlingerSaylor:2018bnu} to which they can be directly compared.

@@

The main contamination is expected to be 
from the exclusive production of a $\pi^0$, the final state of which is very similar to DVCS with 
only an extra photon. In such an event, if one of the photons is produced at low energy, it is 
easy to confuse this process with single photon production. In order to estimate the contribution from
this channel in the data, we measure the exclusive $\pi^0$ production in the same way as DVCS, with 
a series of exclusivity cuts, completed by a selection cut on the invariant mass of the two photons
to match the $\pi^0$ mass. The events obtained for the coherent and incoherent
channels are shown in Fig. \ref{fig:CohPi0Simul} and \ref{fig:InCohPi0Simul}, respectively. From 
this sample, we elaborated an event generator, which after being processed
in the simulation of the detectors gives the red histograms of Fig. \ref{fig:CohPi0Simul} and 
\ref{fig:InCohPi0Simul}. To correct the experimental data, we estimate the number of single photon 
events coming from the exclusive $\pi^0$ production as:
\begin{equation}
	N_{1\gamma,\pi^0}^{Exp} = \frac{N_{1\gamma,\pi^0}^{Sim}}{N_{2\gamma,\pi^0}^{Sim}} \times N_{2\gamma,\pi^0}^{Exp},\\
\end{equation}
where $N_{1\gamma,\pi^0}^{Sim}$ is the number of simulated exclusive $\pi^0$ mistaken for DVCS events,
$N_{2\gamma,\pi^0}^{Sim}$ the number of simulated exclusive $\pi^0$ fully reconstructed and $N_{2\gamma,\pi^0}^{Exp}$
the number of experimentally measured exclusive $\pi^0$. This number can then be subtracted from
the experimentally measured number of DVCS events ($N_{DVCS}^{Exp}$) to get the corrected result: 
\begin{equation}
	N_{DVCS}^{Corr} = N_{DVCS}^{Exp} - N_{1\gamma,\pi^0}^{Exp}.\\
\end{equation}

@@

The $\pi^0$ contamination was found to be 2 to 4\% in the coherent channel and 3 to 17\% in the incoherent
channel (with variations between kinematic bins). To make the correction on the DVCS BSA, we assume that
the exclusive $\pi^0$ production has no such asymmetry. This has been checked with the exclusive 
$\pi^0$ production data, for which no significant level of BSA has been measured.

@@

\subsection{Systematic uncertainties}

To evaluate the systematic error of the measurement, we performed several specialized studies. We 
evaluated the impact of changing the exclusivity selection cuts by varying them from 1 to 5 $\sigma$. 
We evaluated the impact of changing the binning in $\phi$ on the extraction of the beam-spin asymmetry
at 90°. The precision with which we know the beam polarization is estimated based on the spread 
of its measurements during the complete run period. We also checked different methods to make the 
simulation of the exclusive $\pi^0$ production to evaluate the 
possible bias introduced by this correction. As radiative corrections are expected to be small
for this process, we did not apply them, but associated an error of their expected value.
These uncertainties are summarized in Tab. \ref{Table:systematic_uncertainties}, with their respective 
evaluated values. They are added quadratically for the total systematic error presented in the results.

@@

%TODO add sentence on recent theory and one on ALERT
